---
title: "Asthma Equity Explorer Report"
params:
  geo_name: NA
  num_geo: NA
  factor: NA
  map_path: NA
  map_rank_path: NA
  scatter_path: NA
  scatter_rank_path: NA
  data_table: NA
  hist_factor: NA
  hist_asthma: NA
  factor_units: NA
  geo_unit: NA
  factor_def: NA
  interpret_info: NA
  dat_info: NA
  factor_og_name: NA
  selection_sentence: NA
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(ggplot2)
library(stringr)

report_date <- Sys.time()

geo_title <- params$geo_name
num_geo <- params$num_geo
factor <- params$factor
map <- params$map_path
map_rank <- params$map_rank_path
scatter <- params$scatter_path
scatter_rank <- params$scatter_rank_path
data_table <- params$data_table
factor_units <- params$factor_units
geo_unit <- params$geo_unit
factor_def <- params$factor_def
interpret_info <- params$interpret_info
dat_info <- params$dat_info
selection_sentence <- params$selection_sentence
asthma_available <- params$asthma_available

geo_unit_singular <- if (geo_unit == "counties") {
  "county"
} else { "tract"}

colnames(data_table) <- str_remove_all(colnames(data_table), "<br>") %>%
  str_trim()

factor_header <- colnames(data_table)[
  stri_detect_fixed(tolower(str_trim(colnames(data_table))), 
          tolower(str_trim(factor)))
  ]
asthma_header <- colnames(data_table)[
  str_detect(tolower(colnames(data_table)), 
          "asthma")
  ]


```

The Asthma Equity Explorer is an interactive tool intended to help asthma programs, researchers, and community members understand the relationships between the Social Determinants of Health (SDOH) and asthma prevalence within our communities.

```{r message=FALSE, warning=FALSE, include=FALSE}
formatted_def = str_replace_all(factor_def$Definition[1], "<br>", "\n \n \n>    ")
```


## Report Summary

Report generated on: `r report_date`

`r selection_sentence`

`r str_trim(factor)`:

>    *Group:* `r factor_def$Group[1]`
    
>    *Source:* `r factor_def$Source[1]`

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

cat(paste0(">    *Definition:* ", formatted_def))

```

Model Outcome:

> `r interpret_info`

## Figures

The figures shown below are meant to serve as a visual tool for understanding the geographic and population distributions of `r str_trim(factor)` and Asthma Prevalence, as well as any trends between them. They serve to visually describe the relationship between these two factors, and can help illuminate areas where there are high or low associations between them.


![Map - `r str_trim(factor)`](`r map`)

![Map - Rank](`r map_rank`)

The `r str_trim(factor)` map shows **`r str_trim(factor)` (`r factor_units`)** for all `r geo_unit` in `r geo_title` as a color gradient overlay. **Asthma Prevalence (percent)** is shown as grey circles over the center of all `r geo_unit`, where the larger, darker circles represent higher asthma prevalence and the smaller, lighter circles represent lower asthma prevalence. The Rank map shows the the combined rank between asthma prevalence and `r str_trim(factor)` for all `r geo_unit` in `r geo_title` as a red gradient overlay, where the darker red corresponds to a higher rank (closer to 1) and a lighter color corresponds to a lower rank. See the Data Table section for more details on how rank is defined.


![Scatter Plot - `r str_trim(factor)`](`r scatter`)

![Scatter Plot - Rank](`r scatter_rank`)

The `r str_trim(factor)` scatter plot shows **`r str_trim(factor)` (`r factor_units`)** versus **Asthma Prevalence (percent)** for all `r geo_unit` in `r geo_title`, where each `r geo_unit_singular` is represented as a colored circle. The size of the circle corresponds to the population density (population per square mile) in that `r geo_unit_singular`, where larger circles represent higher population density and smaller circles represent lower population density. The dashed line bisecting the plot represents the line of best fit resulting from a linear regression of Asthma Prevalence (response variable) based on `r str_trim(factor)` (explanatory variable) for `r geo_unit` in `r geo_title`. The Rank scatter plot shows the combined rank between asthma prevalence and `r str_trim(factor)` for all `r geo_unit` in `r geo_title` as a red gradient overlay, where the darker red corresponds to a higher rank (closer to 1) and a lighter color corresponds to a lower rank. See the Data Table section for more details on how rank is defined.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Summary Statistics for factor and asthma prevalence
factor_dat <- data_table[[factor_header]]
asthma_dat <- data_table[[asthma_header]]

# Function to create a data frame of summary statistics
create_metrics <- function(dat, name) {
  
  # Calculate information
  dat_mean = mean(dat, na.rm = T)
  dat_median = median(dat, na.rm = T)
  dat_max = max(dat, na.rm = T)
  dat_min = min(dat, na.rm = T)
  dat_quantiles = quantile(dat, na.rm = T)
  names(dat_quantiles) <- NULL
  
  # Put together
  summary_dat <- c(
      "Min" = dat_min,
      "1st Qu." = dat_quantiles[2],
      "Median" = dat_median,
      "Mean" = dat_mean,
      "3rd Qu." = dat_quantiles[4],
      "Max" = dat_max
    )
  
  # Convert to data frame
  final_dat <- data.frame(
      name = summary_dat,
      check.names = FALSE
    ) %>% 
    dplyr::mutate_all(~round(., 3))
  
  # Set row and column names
  colnames(final_dat) <- name
  row.names(final_dat) <- names(summary_dat)
  
  return(
    final_dat %>% 
    tibble::rownames_to_column(var = "Metric")
  )
}

# Convert to flextable
bind_cols(
    create_metrics(factor_dat, factor_header),
    create_metrics(asthma_dat, asthma_header) %>% dplyr::select(-Metric)
  ) %>%
  flextable::flextable() %>%
  flextable::width(
    width = 3.25,
    unit = "cm"
  ) %>%
  flextable::align(
    align = "center",
    part = "all"
  ) %>%
  flextable::bold(
    bold = TRUE,
    part = "all",
    j = 1) %>%
  flextable::align(
    align = "right",
    part = "all",
    j = 1
  ) %>%
  flextable::set_caption("Summary Statistics")
```

The summary statistics provide high level characteristics of the distributions of Asthma Prevalence and `r str_trim(factor)` for all `r geo_unit` in `r geo_title`. The mean Asthma Prevalence (percent) is `r round(mean(asthma_dat, na.rm = T), 3)` and the mean `r str_trim(factor)` (`r factor_units`) is `r round(mean(factor_dat, na.rm = T), 3)`. The histograms below, along with their overlaid density curves, provide further visualizations of these distributions, giving the shape and size of the data.

![`r factor` Distribution](`r params$hist_factor`)

![Asthma Prevalence Distribution](`r params$hist_asthma`)

\pagebreak

## Data Table {#table_section}

```{r message=FALSE, warning=FALSE, include=FALSE}
# assigns order based on variable
cov_order <- dat_info[dat_info$Colname == params$factor_og_name,"Rank.Cardinality"]

cov_pos_neg <- 
  if (cov_order == "increasing") {
    "negative"
  } else {
    "positive"
  }

cov_order <- 
  if (cov_order == "increasing") {
    "decreasing"
  } else {
    "increasing"
  }
```


The following table lists the Asthma Prevalence (percent) and `r str_trim(factor)` (`r factor_units`) for every `r geo_unit_singular` in `r geo_title`. The table is sorted by the Rank score. The Rank is determined by first taking the rank of Asthma Prevalence for all geographic units, then taking the rank of `r str_trim(factor)` for all geographic units (since higher `r str_trim(factor)` corresponds to a `r cov_pos_neg` outcome, the rank is in _`r cov_order`_ order), then averaging them (using the mean), and then taking the rank of the average. The population size of the geographic unit is used to break ties. Note, Race/Ethnicity factors aren't ranked, their corresponding geographies are ranked solely by their Asthma Prevalence.

<br>


```{r echo=FALSE, message=FALSE, warning=FALSE}

# Trim whitespace
# colnames(data_table) <- str_trim(colnames(data_table))
# 
# # Identify columns that we need to append units to
# cols_to_change <- c(which(str_detect(colnames(data_table), "Asthma")),
#                     which(str_detect(colnames(data_table), str_trim(factor))))
# 
# # Append units
# colnames(data_table)[cols_to_change] <-
#   paste0(colnames(data_table)[cols_to_change], c(" (percent)", paste0(" (", factor_units, ")")))

# The GEOID number an empty column name, which is not
# allowed
if (any(colnames(data_table) == "")) {
  colnames(data_table)[which(colnames(data_table) == "")] <- "GEOID"
}

# Print table
data_table %>%
  arrange(Rank) %>%
  flextable::flextable() %>%
  flextable::width(
    width = 3.25,
    unit = "cm"
  ) %>%
  flextable::align(
    align = "center",
    part = "header"
  ) %>%
  flextable::bold(
    bold = TRUE,
    part = "header")

```

