---
title: "Asthma Equity Explorer Report"
params:
  geo_name: NA
  num_geo: NA
  factor: NA
  map_path: NA
  data_table: NA
  hist_factor: NA
  factor_units: NA
  geo_unit: NA
  factor_def: NA
  interpret_info: NA
  dat_info: NA
  factor_og_name: NA
  selection_sentence: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(ggplot2)
library(stringr)

report_date <- Sys.time()

geo_title <- params$geo_name
num_geo <- params$num_geo
factor <- params$factor
data_table <- params$data_table
factor_units <- params$factor_units
geo_unit <- params$geo_unit
factor_def <- params$factor_def
interpret_info <- params$interpret_info
dat_info <- params$dat_info
selection_sentence <- params$selection_sentence
map <- params$map_path

geo_unit_singular <- if (geo_unit == "counties") {
  "county"
} else { "tract"}

colnames(data_table) <- str_remove_all(colnames(data_table), "<br>") %>%
  str_trim()

factor_header <- colnames(data_table)[
  stri_detect_fixed(tolower(str_trim(colnames(data_table))), 
          tolower(str_trim(factor)))
  ]
asthma_header <- colnames(data_table)[
  str_detect(tolower(colnames(data_table)), 
          "asthma")
  ]


```

The Asthma Equity Explorer is an interactive tool intended to help asthma programs, researchers, and community members understand the relationships between the Social Determinants of Health (SDOH) and asthma prevalence within our communities.

```{r message=FALSE, warning=FALSE, include=FALSE}
formatted_def = str_replace_all(factor_def$Definition[1], "<br>", "\n \n \n>    ")
```

## Report Summary

Report generated on: `r report_date`

`r selection_sentence`

`r str_trim(factor)`:

> *Group:* `r factor_def$Group[1]`

> *Source:* `r factor_def$Source[1]`

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

cat(paste0(">    *Definition:* ", formatted_def))

```

Model Outcome:

> `r interpret_info`

See the [Current Release Notes](https://www.cdc.gov/places/help/data-notes/index.html) from CDC's PLACES data for more information on which locations are missing data.

## Figures

The figures shown below are meant to serve as a visual tool for understanding the geographic and population distributions of `r str_trim(factor)`.

![Map - `r str_trim(factor)`](`r map`)

The `r str_trim(factor)` map shows **`r str_trim(factor)` (`r factor_units`)** for all `r geo_unit` in `r geo_title` as a color gradient overlay. Since Asthma Prevalence data isn't available at this geographic level, only `r str_trim(factor)` is shown.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Summary Statistics for factor and asthma prevalence
factor_dat <- data_table[[factor_header]]

# Function to create a data frame of summary statistics
create_metrics <- function(dat, name) {
  
  # Calculate information
  dat_mean = mean(dat, na.rm = T)
  dat_median = median(dat, na.rm = T)
  dat_max = max(dat, na.rm = T)
  dat_min = min(dat, na.rm = T)
  dat_quantiles = quantile(dat, na.rm = T)
  names(dat_quantiles) <- NULL
  
  # Put together
  summary_dat <- c(
      "Min" = dat_min,
      "1st Qu." = dat_quantiles[2],
      "Median" = dat_median,
      "Mean" = dat_mean,
      "3rd Qu." = dat_quantiles[4],
      "Max" = dat_max
    )
  
  # Convert to data frame
  final_dat <- data.frame(
      name = summary_dat,
      check.names = FALSE
    ) %>% 
    dplyr::mutate_all(~round(., 3))
  
  # Set row and column names
  colnames(final_dat) <- name
  row.names(final_dat) <- names(summary_dat)
  
  return(
    final_dat %>% 
    tibble::rownames_to_column(var = "Metric")
  )
}

# Convert to flextable
create_metrics(factor_dat, factor_header) %>%
  flextable::flextable() %>%
  flextable::width(
    width = 3.25,
    unit = "cm"
  ) %>%
  flextable::align(
    align = "center",
    part = "all"
  ) %>%
  flextable::bold(
    bold = TRUE,
    part = "all",
    j = 1) %>%
  flextable::align(
    align = "right",
    part = "all",
    j = 1
  ) %>%
  flextable::set_caption("Summary Statistics")
```

The summary statistics provide high level characteristics of the distributions of `r str_trim(factor)` for all `r geo_unit` in `r geo_title`. The mean `r str_trim(factor)` (`r factor_units`) is `r round(mean(factor_dat, na.rm = T), 3)`. The histogram below, along with its overlaid density curve, provides a further visualization of this distribution, giving the shape and size of the data.

![`r factor` Distribution](`r params$hist_factor`)

\pagebreak

## Data Table {#table_section}

```{r message=FALSE, warning=FALSE, include=FALSE}
# assigns order based on variable
cov_order <- dat_info[dat_info$Colname == params$factor_og_name,"Rank.Cardinality"]

cov_pos_neg <- 
  if (cov_order == "increasing") {
    "negative"
  } else {
    "positive"
  }

cov_order <- 
  if (cov_order == "increasing") {
    "decreasing"
  } else {
    "increasing"
  }
```

The following table lists the Asthma Prevalence (percent) and `r str_trim(factor)` (`r factor_units`) for every `r geo_unit_singular` in `r geo_title`. The table is sorted by the Rank score. The Rank is traditionally determined by first taking the rank of Asthma Prevalence for all geographic units, then taking the rank of `r str_trim(factor)` for all geographic units (since higher `r str_trim(factor)` corresponds to a `r cov_pos_neg` outcome, the rank is in *`r cov_order`* order), then averaging them (using the mean), and then taking the rank of the average. The population size of the geographic unit is used to break ties. Note, Race/Ethnicity factors aren't ranked, their corresponding geographies are ranked solely by their Asthma Prevalence.

***NOTE:*** *Since there is no Asthma Prevalence data at this geographic level, the rank is determined solely by the factor. If the factor falls under the Race/Ethnicity category, it is not meaningful and therefore every geographic unit gets assigned a value of one.*

<br>

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Trim whitespace
# colnames(data_table) <- str_trim(colnames(data_table))
# 
# # Identify columns that we need to append units to
# cols_to_change <- c(which(str_detect(colnames(data_table), "Asthma")),
#                     which(str_detect(colnames(data_table), str_trim(factor))))
# 
# # Append units
# colnames(data_table)[cols_to_change] <-
#   paste0(colnames(data_table)[cols_to_change], c(" (percent)", paste0(" (", factor_units, ")")))

# The GEOID number an empty column name, which is not
# allowed
if (any(colnames(data_table) == "")) {
  colnames(data_table)[which(colnames(data_table) == "")] <- "GEOID"
}

# Print table
data_table %>%
  arrange(Rank) %>%
  flextable::flextable() %>%
  flextable::width(
    width = 3.25,
    unit = "cm"
  ) %>%
  flextable::align(
    align = "center",
    part = "header"
  ) %>%
  flextable::bold(
    bold = TRUE,
    part = "header")

```
